---
title: "09.20 hw1"
author: "Guangling Xu"
date: "9/16/2019"
output: github_document
---

## Set up
```{r set up, include=FALSE}
library(tidyverse)
##code will be executed but not displayed; results are not included
```


## Problem 1

```{r problem1,  warning = FALSE}
df_1 = tibble(
   samp = rnorm(8),  
   # a random sample of size 8 from a standard Normal distribution  
   vec_logical = samp > 0, 
   #a logical vector indicating whether elements of the sample are greater than 0
   vec_char = c("This is R!"), 
   #a character vector of length 8
   vec_factor = factor(c("sun","rain","wind","wind","sun","rain","sun","rain")), 
   #a factor vector of length 8, with 3 different factor “levels”
  
)
```
## Take mean of each variable in the dataframe
* Mean of random sample:
```{r} 
mean(df_1$samp) ## works, data is numerical
```
* Mean of logical vector:
```{r}
mean(df_1$vec_logical) ## works, TRUE = 1 and FALSE = 0
```
* Mean of character vector: 
```{r}
mean(df_1$vec_char) ## does not work,data is not numerical
```
* Mean of factor vector: 
```{r}
mean(df_1$vec_factor) ## does not work,data is not numerical
```




```{r change data type,results ='hide', warning=FALSE}
as.numeric(df_1$samp) 
## works, random sample is composed of numeric data
as.numeric(df_1$vec_logical) 
## works, TRUE = 1 and FALSE = 0
as.numeric(df_1$vec_char) 
## characters cannot be transformed into numeric data
as.numeric(df_1$vec_factor) 
## The levels of a factor are stored as character data type, but transforming factor into numeric data is meaningless
```

## Convert data type

* convert the logical vector to numeric,and multiply the random sample by the result
```{r, warning = FALSE}
(as.numeric(df_1$vec_logical)) *(df_1$samp) 
```

* convert the logical vector to logical,and multiply the random sample by the result
```{r}
(as.factor(df_1$vec_logical)) * (df_1$samp) ##factors cannot be multiplied
```

* convert the logical vector to a factor and then convert the result to numeric, and multiply the random sample by the result
```{r}
(as.numeric(as.factor(df_1$vec_logical))) * (df_1$samp)
## numerics are 2,1,2,2,1,1,2,1
```

## Section 2

```{r problem2}
set.seed(1234)
df_2 = tibble(
  
  x = rnorm(500), ## a random sample of size 500 from a standard Normal distribution
  y = rnorm(500), ## a random sample of size 500 from a standard Normal distribution
  vec_logical1 = x + y > 1, ## a logical vector indicating whether x + y > 1
  vec_numerical1 = as.numeric(vec_logical1), ## a numeric vector created by coercing the above logical vector
  vec_factor1 = as.factor(vec_logical1), ## A factor vector created by coercing the above logical vector
  
)
```
## Calculate the size of the dataset and the mean, median, standard deviation of x and the proportion of cases for which x + y > 1

* Number of rows:
```{r}
nrow(df_2)
```
* Number of colums: 
```{r}
ncol(df_2)
```
* Mean of x: 
```{r}
mean(df_2$x)
```
* Median of x:
```{r}
median(df_2$x)
```
* Standard diviation of x:  
```{r}
sd(df_2$x, na.rm = FALSE)
```
* proportion of cases for which x + y > 1: 
```{r}
sum(df_2$vec_logical1 == TRUE) / (sum(df_2$vec_logical1 == FALSE) + sum(df_2$vec_logical1 == TRUE))
```


# Make a scatterplot
```{r}
ggplot(df_2, aes(x = x,y = y,color = df_2$vec_logical1,)) + geom_point()+
  labs(title="y vs x;color points using the logical variables", x="random(500)", y = "random(500)")+
  theme_classic()  
ggsave("Scatter plot1 for hw1.pdf")
ggplot(df_2, aes(x = x,y = y,color = df_2$vec_numerical1,)) + geom_point()+
  labs(title="y vs x;color points using the numeric variables",
x="random(500)", y = "random(500)")+
  theme_classic()  
ggsave("Scatter plot2 for hw1.pdf")
ggplot(df_2, aes(x = x,y = y,color = df_2$vec_factor1,)) + geom_point()+
  labs(title="y vs x;color points using the factor variables",
       x="random(500)", y = "random(500)")+
  theme_classic() 
ggsave("Scatter plot3 for hw1.pdf")

```

* When the data type is logical, it shows two colors indicating the "TRUE" and "FALSE" logical outcomes.

* When the data type is numerical, the scale on the right side using different blue colors to differenciate numbers. 
since TRUE=1, FALSE=0, the color points have only two colors indicating 1 and 0.

* When the data types is factor, the color points represent "TRUE" and "FALSE" factor in blue and red.





